---
title: "Data Scientist Jobs Analysis"
author: "Shang Liang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(tidyverse)
```


## Data Cleaning


### Exclusions

Not all results from LinkedIn's search are accurate. Besides "data scientist" jobs, other jobs like `Software Engineer` are listed. Possibly these role are under the same department or the job description contains the keyword. Although roles like `Machine Learning Engineer` or `Data Engineer` are more related to `Data Scientist`, this analysis will focus on pure `Data Scientist` jobs. To enforce it, a filter on the job title is apply to ensure that it contain `Data Scientist` keyword. 

Job listings from China are mostly written in Chinese. For this analysis, they will be excluded. 
```{r include=F}
raw<-read_csv('./scraper/data/jobs.csv')
df <- raw %>%
  filter(grepl('data scientist', title, ignore.case = T),jobLocation.address.addressCountry!='CN') %>%
  select(
    title,
    company = identifier.name,
    datePosted,
    validThrough,
    employmentType,
    industry,
    country = jobLocation.address.addressCountry,
    city = jobLocation.address.addressLocality,
    education = educationRequirements.credentialCategory,
    salaryCurrency = baseSalary.currency,
    salaryMax = baseSalary.value.maxValue,
    salaryMin = baseSalary.value.minValue,
  )
```
b
### Adding Seniority

The seniority of a job will be derived from the job title. Based on the observation of the data, a keyword near to `Data Scientist` normally indicates the level. The following 4 patterns are observer (*X* denoting the job level keyword).

* level *X* data scientist
* data scientist (level *X*)
* *X* data scientist
* data scientist *X* 

The most common levels are  

* Junior Data Scientist
* Data Scientist
* Senior Data Scientist
* Staff Data Scientist

There are a few Lead and Principal titles and they are grouped together as `Staff` level. The level extraction also handles abbreviations like snr, sr for Senior and jr for Junior.  

A different notation is Data Scientist I/1, II/2, III/3, IV/4. They are mapped to the Junior to Staff level accordingly. 

```{r}

cleanLevel <- function(l) {
  if (l %in% c("senior", "sr", "3", "iii")) {
    return("senior")
  } else if (l %in% c("junior","jr", "1", "i", "associate", "asso")) {
    return("junior")
  } else if (l %in% c("staff", "principal", "lead", "4", "iv", "v", "5")) {
    return("staff")
  }
  NA
}
deriveLevel<-function(s) {
  s<-str_to_lower(s)
  # Remove special characters to have better word boundaries
  s<-str_replace_all(s,"[()-:/]"," ")
  # keyword check on "senior", "junior", "staff", "principal"
  # some posting has multiple keywords like "senior/staff", will take the first occurrence
  p<-str_extract(s,"\\bsenior\\b|\\bjunior\\b|staff\\b|principal\\b")
  
  if(!is.na(p)){
    # Found a common keyword
   return(cleanLevel(p))
  }
  
  # match pattern `scientist 1,2,3,4,i,ii,iii,iv` and `level 1,2,3,4,i,ii,iii,iv`
  p <- str_extract(s, "(data scientist)|level\\s+[iv0-9]+")
  if(!is.na(p)){
    arr<-str_split(p," ",simplify = T)
    return(cleanLevel(arr[length(arr)]))
  }
  
}
df$level<-sapply(df$title,deriveLevel)
# arrange(as.data.frame(table(df$level)),desc(Freq))
```

## Jobs by company
```{r}
df %>%
  filter(country == 'SG') %>%
  group_by(company) %>%
  summarise(openings=n()) %>%
  arrange(-openings)

```
```{r}
df %>% filter(grepl('manager',title,ignore.case=T)) %>%
  arrange(validThrough)
```
