---
title: "Data Scientist Jobs Analysis"
author: "Shang Liang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(tidyverse)
library(kable)
```


## Data Cleaning


### Exclusions

Not all results from LinkedIn's search are accurate. Besides "data scientist" jobs, other jobs like `Software Engineer` are listed. Possibly these role are under the same department or the job description contains the keyword. Although roles like `Machine Learning Engineer` or `Data Engineer` are more related to `Data Scientist`, this analysis will focus on pure `Data Scientist` jobs. To enforce it, a filter on the job title is apply to ensure that it contain `Data Scientist` keyword. 

Job listings from China are mostly written in Chinese. For this analysis, they will be excluded. 
```{r include=F}
raw<-read_csv('./scraper/data/jobs.csv')
df <- raw %>%
  filter(grepl('data scientist', title, ignore.case = T),jobLocation.address.addressCountry!='CN') %>%
  select(
    title,
    company = identifier.name,
    datePosted,
    validThrough,
    employmentType,
    industry,
    country = jobLocation.address.addressCountry,
    city = jobLocation.address.addressLocality,
    education = educationRequirements.credentialCategory,
    salaryCurrency = baseSalary.currency,
    salaryMax = baseSalary.value.maxValue,
    salaryMin = baseSalary.value.minValue,
  )
```

### Adding Seniority

The seniority of a job will be derived from the job title. Based on the observation of the data, a keyword adjacent to `Data Scientist` normally indicates the level, e.g. Senior/Sr., Lead, Principle, Associate/Asso. 

```{r}
deriveLevel<-function(s) {
  s<-str_to_lower(s)
  s<-str_replace_all(s,"[()-:]","")
  level=""
  # try to match "senior level", "junior level" etc
  p <- str_extract(s, "[a-zA-Z0-9\\.]+\\s+level")
  if(is.na(p)) {
    # check if any useful word in front of "data scientist"
    p <- str_extract(s, "[a-zA-Z0-9\\.]+\\s+data scientist")
    # if nothing in front is useful, try the word behind
    if (is.na(p)) {
      p <- str_extract(s, "[dD]ata [sS]cientist\\s+[a-zA-Z0-9\\.]+")
      if (!is.na(p)) {
        level <- substr(p, 15, str_length(p))
      }
    } else{
      level <- substr(p, 0, str_length(p) - 15)
    }
  }else{
    level<-substr(p,0,str_length(p)-6)
  }
  
  if (level %in% c("2", "ii")) {
    level = ""
  } else if (level %in% c("sr", "3", "ii")) {
    level <- "senior"
  } else if (level %in% c("jr", "1", "i", "associate", "asso")) {
    level <- "junior"
  } else if (level %in% c("principle", "lead", "4", "iv", "v", "5")) {
    level <- "staff"
  }
  if (!(level %in% c("junior", "senior", "staff"))) {
    level <- ""
  }
  level
}
df$level<-sapply(df$title,deriveLevel)
arrange(as.data.frame(table(df$level)),desc(Freq))
```

## Jobs by company
```{r}
df %>%
  filter(country == 'SG') %>%
  group_by(company) %>%
  summarise(openings=n()) %>%
  arrange(-openings)

```
```{r}
df %>% filter(grepl('manager',title,ignore.case=T),country=='SG') %>%
  arrange(validThrough)
```
